---
title: "covr thinks I'm CRAN!"
author: "David Fong"
date: 2020-09-10T00:00:00Z
categories: ["R"]
image:
  caption: <span>Photo by <a href="https://unsplash.com/@chrisliverani?utm_source=unsplash&amp;utm_medium=referral&amp;utm_content=creditCopyText">Chris Liverani</a> on <a href="https://unsplash.com/s/photos/exam?utm_source=unsplash&amp;utm_medium=referral&amp;utm_content=creditCopyText">Unsplash</a></span>
  focal_point: Smart
aliases:
tags: ["R", "covr", "testthat", "NOT_CRAN"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

[`testthat`](https://r-pkgs.org/tests.html) is a great tool to test that R package
code actually does what it is supposed to do. And keep doing the right thing
after multiple package updates, edits and 'fixes'.

[`covr`](https://github.com/r-lib/covr) [helps show what part of the codebase
is actually tested with `testthat`](https://codingclubuc3m.rbind.io/post/2019-11-26/).

It is possible to [use testthat with shiny](https://github.com/rfaelens/exampleShinyTest).

One annoying problem I had is that `devtools::test()` might work : 

```{r eval=FALSE}
> devtools::test()
Loading DTedit
Testing DTedit
✓ |  OK F W S | Context
✓ |   1       | callbacks_actionButtons [12.5 s]
✓ |   1       | dtedit_demo [10.0 s]
✓ |   1       | dtedit_demo_modular [5.2 s]
✓ |   5       | error_test [6.5 s]
✓ |   1       | fileInput_modular [12.2 s]
✓ |   1       | password [6.9 s]
✓ |   1       | reactive [2.7 s]
✓ |   1       | reactive_demo [2.7 s]
✓ |   1       | selectInputReactive [8.7 s]
✓ |   1       | selectInputReactive_demo [13.6 s]
✓ |   1       | simple [13.6 s]
✓ |   1       | simple_modular [13.7 s]

══ Results ═════════════════════════════════════════════════════════════════════
Duration: 113.0 s

OK:       16
Failed:   0
Warnings: 0
Skipped:  0
```

<p>
but `testthat::test_package()` skips all the tests :

```{r eval = FALSE}
> testthat::test_package("DTedit", quiet = FALSE)
══ testthat results  ═══════════════════════════════════════════════════════════
[ OK: 0 | SKIPPED: 12 | WARNINGS: 0 | FAILED: 0 ]
```

<p>

with similar results from `covr::codecov()` :

```{r eval=FALSE}
> covr::codecov(quiet = FALSE)
* installing *source* package ‘DTedit’ ...
** R
** inst
** tests
** byte-compile and prepare package for lazy loading
** help
*** installing help indices
  converting help for package ‘DTedit’
    dtedit                                  example
    dteditmodUI                             example
*** copying figures
** building package indices
** installing vignettes
** testing if installed package can be loaded
* DONE (DTedit)
Running specific tests for package ‘DTedit’
  Running ‘testthat.R’
DTedit Coverage: 0.00%
R/dtedit_demo.R: 0.00%
R/dtedit_test.R: 0.00%
R/dtedit.R: 0.00%
```

<p>
ouch! no coverage!

The problem is that, strangely, [the default setting of various R packages is to
think that testing is occurring on a CRAN computer](https://github.com/r-lib/covr/issues/314)!

So tests will be skipped if `testthat::skip_on_cran()` is used in the tests.

There a few possible solutions, but one solution 
[described by Randy Lai](https://github.com/r-lib/covr/issues/314) is to use
`withr::with_envvar` :

```{r eval=FALSE}
> withr::with_envvar(c("NOT_CRAN" = "true"), covr::codecov(quiet = FALSE))
* installing *source* package ‘DTedit’ ...
** R
** inst
** tests
** byte-compile and prepare package for lazy loading
** help
*** installing help indices
  converting help for package ‘DTedit’
    dtedit                                  example
    dteditmodUI                             example
*** copying figures
** building package indices
** installing vignettes
** testing if installed package can be loaded
* DONE (DTedit)
Running specific tests for package ‘DTedit’
  Running ‘testthat.R’
DTedit Coverage: 95.07%
R/dtedit.R: 94.99%
R/dtedit_test.R: 95.04%
R/dtedit_demo.R: 100.00%
$uploaded
[1] TRUE
```