---
title: "Microsoft R Open 4.0.2 and openSuSE"
author: "David Fong"
date: 2020-09-26T00:00:00Z
categories: ["R"]
image:
  caption: <span>Photo by <a href="https://unsplash.com/@musiime?utm_source=unsplash&amp;utm_medium=referral&amp;utm_content=creditCopyText">Musiime P. Muramura</a> on <a href="https://unsplash.com/s/photos/gecko?utm_source=unsplash&amp;utm_medium=referral&amp;utm_content=creditCopyText">Unsplash</a></span>
  focal_point: Smart
aliases:
tags: ["R", "Microsoft R Open", "4.0.2", "openSuSE Leap 15.2"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

One reason for my sometimes growing respect for Microsoft is that they support
(just) [Microsoft R Open](https://mran.microsoft.com/). Back in May 2019 they released
[version 3.5.3](https://blog.revolutionanalytics.com/2019/05/availability-of-microsoft-r-open-352-and-353.html). It was a long time coming, but Microsoft just released 
[Microsoft R Open 4.0.2](https://blog.revolutionanalytics.com/2020/09/mro-402-available.html).

In the nick-of-time, really. [Theresa Graefe's RTutor tutorial on synthetic models](https://github.com/TheresaGraefe/RTutorCarbonTax) appears to rely on at least version 3.6 of R.

## Microsoft R Open 4.0.2 and openSuSE Leap 15.2

I mainly use Microsoft R on Windows, but Linux has once again become my 'daily driver'
for home use. Microsoft R Open does [support SuSE Linux Enterprise SLES](https://mran.microsoft.com/documents/rro/installation) installation. To install on openSuSE Leap 15.2, I needed to change line 109 of the `install.sh` to:

```{r eval=FALSE}
  elif  [ "$OS_NAME" == "openSUSE" ]; then
```

## Microsoft R Open CRAN snapshots

Normally, I think snapshots are a good idea to ease maintenance (though it would be more usable if Microsoft had a more regular update cycle...).

For some reason, the default CRAN mirror isn't quite working the way it should be.

`RevoUtils::getRevoRepos()` returns `"https://mran.microsoft.com/snapshot/2020-07-16"`.

but...

```{r eval=FALSE}
> install.packages("tidyverse")
Warning in install.packages :  
  unable to access index for repository https://mran.microsoft.com/snapshot/2020-07-16/src/contrib:  
  Line starting '<head><title>Documen ...' is malformed!
```

The problem can be 'fixed' manually after R starts with the commands:

```{r eval=FALSE}
r <- getOption("repos")
r["CRAN"] <- "https://cran.microsoft.com/snapshot/2020-07-16"
options(repos = r)
```

but a more permanent fix is by modifying line 23 of `/opt/microsoft/ropen/4.0.2/lib64/R/etc/Rprofile.site`.

```{r eval=FALSE}
 r["CRAN"] <- "https://cran.microsoft.com/snapshot/2020-07-16" # RevoUtils::getRevoRepos()
```