---
title: "Kenya Spur Afrika 2023 - Lord of the Dance!"
authors: ["David Fong"]
date: '2023-01-11T11:25:00Z'
categories: Christianity
aliases: null
tags:
- Christianity
- Kenya
- Kibera
- Kisumu
- Spur Afrika
- Spur Afrika 2023
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

> "Praise [God] for his mighty deeds;  
praise him according to the abundance of his greatness.  
Praise him with blast of horn;  
praise him with harp and lyre.  
Praise him with tambourine and dancing;  
praise him with strings and flute.  
Praise him with sounding cymbals;  
praise him with clashing cymbals."
>   
*Psalm 150:2-5*

A wonderful feature of Kenyan life is community dancing and being ... loud!

Be it at the Spur Afrika picnic day  in Kibera/Nairobi (first video), or the welcome to Holo/Kisumu that the Spur Afrika children gave Rosalie and myself yesterday (second video), there seems to be no end to the dancing and singing!

Church services that we've attended are like that as well. Lots of dancing and movement (by those on stage. and the congregation). And almost deafening music!

Churches have, over centuries, been at the forefront of loud music and noises, be it the pipe organ (technological marvels of the day) and pealing bells.

Something about public praise, worship and joy the Kenyans haven't yet forgotten!

[Spur Afrika trip 2023 posts](/spurafrika2023/)

```{r echo=FALSE}
blogdown::shortcode('video', library = 'true', src = '20230107_team_danceoff.mp4', controls = 'yes')
```

```{r echo=FALSE}
blogdown::shortcode('video', library = 'true', src = '20230109_kisumu_welcome.mp4', controls = 'yes')
```